---
title: "Workplace Experience & Retention Analytics in Nigerian Professional Services"
author: "Hannah Ishola-Oshodi"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    toc-location: left
    code-fold: true
    code-summary: "Show Code"
    self-contained: true
    smooth-scroll: true
execute:
  warning: false
  message: false
  echo: true
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(ggplot2)
library(scales)
library(corrplot)
library(patchwork)
library(broom)
library(effectsize)
library(moments)
library(RColorBrewer)
library(gt)
library(glue)

palette_cat <- c("#1B3A6B","#E8A020","#2E86AB","#A23B72","#4CAF50","#FF6B35")

theme_lbs <- function(base_size = 12) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title       = element_text(face = "bold", size = base_size + 3,
                                      colour = "#1B3A6B", hjust = 0),
      plot.subtitle    = element_text(size = base_size, colour = "#555555", hjust = 0),
      axis.title       = element_text(face = "bold", colour = "#2C2C2C"),
      axis.text        = element_text(colour = "#444444"),
      panel.grid.major = element_line(colour = "#EBEBEB", linewidth = 0.4),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background  = element_rect(fill = "white", colour = NA),
      legend.position  = "bottom",
      plot.margin      = margin(15, 15, 10, 15)
    )
}
theme_set(theme_lbs())
```

```{r}
#| label: load-data

raw <- read.csv("survey_data.csv", stringsAsFactors = FALSE, check.names = FALSE)

names(raw) <- c("timestamp","gender","education","experience","role",
                "firm_size","hours","training","satisfaction","work_life",
                "performance","supervisor","compensation","leaving",
                "leave_reason","recommend","tenure")

df <- raw

df[] <- lapply(df, function(x) if (is.character(x)) trimws(x) else x)
df[] <- lapply(df, function(x) if (is.character(x)) gsub("\u2013", "-", x) else x)

df$satisfaction  <- as.numeric(df$satisfaction)
df$work_life     <- as.numeric(df$work_life)
df$performance   <- as.numeric(df$performance)
df$supervisor    <- as.numeric(df$supervisor)
df$compensation  <- as.numeric(df$compensation)

df$performance[is.na(df$performance)] <- median(df$performance, na.rm = TRUE)
df$training[is.na(df$training) | df$training == ""] <- "None"
df <- df[!is.na(df$experience) & df$experience != "", ]

df$experience <- factor(df$experience,
  levels = c("Less than 2 years","2-5 years","6-10 years","11-15 years","More than 15 years"),
  ordered = TRUE)
df$hours <- factor(df$hours,
  levels = c("Less than 35","35-40","41-50","51-60","More than 60"), ordered = TRUE)
df$training <- factor(df$training,
  levels = c("None","1-2","3-5","6-10","More than 10"), ordered = TRUE)
df$tenure <- factor(df$tenure,
  levels = c("Less than 1 year","1-3 years","4-7 years","8-12 years","More than 12 years"),
  ordered = TRUE)
df$firm_size <- factor(df$firm_size,
  levels = c("1-20","21-50","51-100","101-250","More than 250"), ordered = TRUE)
df$leaving <- factor(df$leaving,
  levels = c("No, I intend to stay","Unsure","Yes, but not actively looking","Yes, actively looking"))
df$recommend <- factor(df$recommend,
  levels = c("Definitely not","Probably not","Not sure","Probably yes","Definitely yes"))
df$gender       <- factor(df$gender)
df$education    <- factor(df$education)
df$role         <- factor(df$role)
df$leave_reason <- factor(df$leave_reason)

df$leaving_binary <- ifelse(df$leaving == "No, I intend to stay", 0L, 1L)
df$experience_num <- as.numeric(df$experience)
df$training_num   <- as.numeric(df$training)
df$hours_num      <- as.numeric(df$hours)
df$tenure_num     <- as.numeric(df$tenure)

n_obs <- nrow(df)
```

---

## 1. Executive Summary

This study investigates workplace experience and employee retention intentions among `r n_obs` professionals in Nigerian legal and professional services organisations. A structured online survey captured demographic information, working conditions, and subjective ratings of satisfaction, work-life balance, performance, supervisor support, and compensation fairness.

Five analytical techniques were applied. **Exploratory Data Analysis** revealed that nearly 73% of respondents express some intention to leave their current organisation within 12 months. **Data Visualisation** exposed systematic patterns in satisfaction across role types and firm sizes. **Hypothesis Testing** confirmed statistically significant differences in job satisfaction across role categories. **Correlation Analysis** found that supervisor support and compensation fairness are the strongest predictors of overall satisfaction. **Logistic Regression** identified compensation fairness and tenure as the most significant predictors of attrition intent.

**Key Recommendation:** Nigerian professional-services organisations should conduct an immediate compensation benchmarking exercise and implement structured supervisor development programmes targeting staff in their first three years of employment.

---

## 2. Professional Disclosure

**Job Title:** Executive Assistant
**Organisation Type:** Legal Services (Law Firm), Lagos, Nigeria

**Why these five techniques?**

**EDA** helps me understand workforce data distributions before drawing conclusions - directly relevant to preparing HR briefings for managing partners. **Visualisation** translates findings into board-ready charts for time-constrained legal partners. **Hypothesis Testing** provides statistical rigour to confirm whether perceived differences between staff groups are real. **Correlation Analysis** identifies which HR investments actually move satisfaction scores. **Logistic Regression** predicts which staff are at highest attrition risk, enabling proactive retention interventions.

---

## 3. Data Collection & Sampling

Data were collected via a structured online questionnaire administered through Google Forms between 28 April 2026 and 1 May 2026. Respondents were recruited through professional WhatsApp networks, LinkedIn, and direct email contact using purposive convenience sampling.

| Parameter | Detail |
|-----------|--------|
| Population | Professionals in Nigerian legal and professional-services firms |
| Sampling method | Purposive convenience sampling |
| Sample size | `r n_obs` usable responses |
| Time period | 28 April - 1 May 2026 |
| Platform | Google Forms (fully anonymous) |

The survey was fully anonymous - no names or email addresses were collected. Participation was entirely voluntary with a consent statement at the top of the form.

---

## 4. Data Description

```{r}
#| label: data-overview

score_vars <- c("satisfaction","work_life","performance","supervisor","compensation")
score_labels <- c("Job Satisfaction","Work-Life Balance","Job Performance",
                  "Supervisor Support","Compensation Fairness")

summary_df <- data.frame(
  Variable = score_labels,
  N        = sapply(score_vars, function(v) sum(!is.na(df[[v]]))),
  Mean     = round(sapply(score_vars, function(v) mean(df[[v]], na.rm=TRUE)), 2),
  Median   = sapply(score_vars, function(v) median(df[[v]], na.rm=TRUE)),
  SD       = round(sapply(score_vars, function(v) sd(df[[v]], na.rm=TRUE)), 2),
  Min      = sapply(score_vars, function(v) min(df[[v]], na.rm=TRUE)),
  Max      = sapply(score_vars, function(v) max(df[[v]], na.rm=TRUE))
)

knitr::kable(summary_df, caption = "Descriptive Statistics - Numeric Rating Variables (1-10 scale)")
```

**Data quality issues addressed:**

- 19 missing values in `training` - treated as "None" (zero sessions)
- 1 missing value in `performance` - replaced with column median
- 1 missing value in `experience` - row excluded

---

## 5. Technique 1 - Exploratory Data Analysis

**Business justification:** Before a managing partner acts on staff welfare data, they need to understand who is in the data and how scores are distributed. EDA makes every subsequent analysis credible.

```{r}
#| label: eda-demographics
#| fig-height: 8
#| fig-width: 10

p1 <- df |> count(gender) |>
  ggplot(aes(x = fct_reorder(gender, n), y = n, fill = gender)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = n), hjust = -0.2, fontface = "bold") +
  scale_fill_manual(values = palette_cat) +
  coord_flip() + labs(title = "Gender", x = NULL, y = "Count") + ylim(0, 55)

p2 <- df |> count(role) |>
  mutate(role = str_wrap(role, 20)) |>
  ggplot(aes(x = fct_reorder(role, n), y = n, fill = role)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = n), hjust = -0.2, fontface = "bold") +
  scale_fill_manual(values = palette_cat) +
  coord_flip() + labs(title = "Role Type", x = NULL, y = "Count") + ylim(0, 30)

p3 <- df |> count(leaving) |>
  ggplot(aes(x = fct_reorder(leaving, n), y = n, fill = leaving)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(n," (",round(n/n_obs*100),"%)")),
            hjust = -0.1, fontface = "bold", size = 3.5) +
  scale_fill_manual(values = c("#4CAF50","#E8A020","#FF6B35","#A23B72")) +
  coord_flip() + labs(title = "Attrition Intention", x = NULL, y = "Count") + ylim(0,40)

p4 <- df |>
  select(satisfaction, work_life, performance, supervisor, compensation) |>
  pivot_longer(everything(), names_to = "var", values_to = "score") |>
  mutate(var = case_when(
    var == "satisfaction" ~ "Satisfaction",
    var == "work_life"    ~ "Work-Life",
    var == "performance"  ~ "Performance",
    var == "supervisor"   ~ "Supervisor",
    var == "compensation" ~ "Compensation",
    TRUE ~ var)) |>
  ggplot(aes(x = var, y = score, fill = var)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  scale_fill_manual(values = palette_cat) +
  scale_y_continuous(breaks = 1:10) +
  labs(title = "Score Distributions", x = NULL, y = "Score (1-10)")

(p1 + p2) / (p3 + p4) +
  plot_annotation(
    title    = "Respondent Profile - Nigerian Professional Services Survey",
    subtitle = "April-May 2026 | n = 100 respondents",
    theme    = theme_lbs()
  )
```

**Interpretation:** The sample is evenly split between male and female professionals. `r round(sum(df$leaving != "No, I intend to stay")/n_obs*100)`% express some intention to leave - a significant talent retention risk. Compensation fairness scores the lowest at a mean of `r round(mean(df$compensation), 1)` out of 10.

---

## 6. Technique 2 - Data Visualisation

**Business justification:** Legal partners are time-constrained. Visualisation translates statistical patterns into decisions. These five charts tell one story: who is most at risk of leaving and why.

```{r}
#| label: viz-story
#| fig-height: 6
#| fig-width: 10

# Chart 1: Attrition by role
df |> count(role, leaving) |>
  group_by(role) |> mutate(pct = n / sum(n)) |> ungroup() |>
  mutate(role = str_wrap(role, 20)) |>
  ggplot(aes(x = role, y = pct, fill = leaving)) +
  geom_col(position = "fill", width = 0.65) +
  scale_fill_manual(values = c("#4CAF50","#A8C8E0","#E8A020","#A23B72"),
                    name = "Intention to Leave") +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Chart 1 - Attrition Intention by Role Type",
       subtitle = "Administrative and IT staff show the highest leaving intentions",
       x = NULL, y = "Proportion") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
#| fig-height: 6
#| fig-width: 10

# Chart 2: Satisfaction by role
df |> group_by(role) |>
  summarise(mean_sat = mean(satisfaction), .groups = "drop") |>
  mutate(role = fct_reorder(str_wrap(role, 22), mean_sat)) |>
  ggplot(aes(x = role, y = mean_sat, fill = role)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = round(mean_sat, 1)), hjust = -0.2, fontface = "bold") +
  scale_fill_manual(values = palette_cat) +
  coord_flip() +
  scale_y_continuous(limits = c(0, 10)) +
  labs(title = "Chart 2 - Mean Job Satisfaction by Role",
       x = NULL, y = "Mean Satisfaction Score (1-10)")
```

```{r}
#| fig-height: 6
#| fig-width: 10

# Chart 3: Reasons for leaving
df |> filter(leave_reason != "I am not considering leaving") |>
  count(leave_reason) |>
  mutate(pct = n / sum(n), leave_reason = fct_reorder(leave_reason, n)) |>
  ggplot(aes(x = leave_reason, y = pct, fill = leave_reason)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = paste0(n," (",percent(pct, accuracy=1),")")),
            hjust = -0.1, fontface = "bold", size = 4) +
  scale_fill_manual(values = palette_cat) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.55)) +
  coord_flip() +
  labs(title = "Chart 3 - Primary Reasons for Considering Leaving",
       x = NULL, y = "Share of Responses")
```

```{r}
#| fig-height: 6
#| fig-width: 10

# Chart 4: Compensation vs Satisfaction scatter
ggplot(df, aes(x = compensation, y = satisfaction, colour = leaving)) +
  geom_jitter(size = 3, alpha = 0.7, width = 0.15, height = 0.15) +
  geom_smooth(aes(group = 1), method = "lm", se = TRUE,
              colour = "#2C2C2C", fill = "#DDDDDD", linewidth = 1) +
  scale_colour_manual(values = c("#4CAF50","#E8A020","#FF6B35","#A23B72"),
                      name = "Intention to Leave") +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(breaks = 1:10) +
  labs(title = "Chart 4 - Compensation Fairness vs Job Satisfaction",
       x = "Compensation Fairness (1-10)", y = "Job Satisfaction (1-10)")
```

```{r}
#| fig-height: 6
#| fig-width: 10

# Chart 5: Satisfaction heatmap
df |> group_by(firm_size, role) |>
  summarise(mean_sat = mean(satisfaction), .groups = "drop") |>
  mutate(role = str_wrap(role, 18)) |>
  ggplot(aes(x = firm_size, y = role, fill = mean_sat)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = round(mean_sat, 1)), fontface = "bold",
            colour = "white", size = 4) +
  scale_fill_gradientn(colours = c("#A23B72","#E8A020","#4CAF50"),
                       limits = c(1,10), name = "Mean Satisfaction") +
  labs(title = "Chart 5 - Satisfaction Heatmap: Role x Firm Size",
       x = "Firm Size", y = NULL) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

**Integrated story:** Administrative and IT staff show the highest attrition risk with the lowest satisfaction. Compensation and better opportunities dominate as leaving reasons. Smaller firms consistently score lower on satisfaction across most roles.

---

## 7. Technique 3 - Hypothesis Testing

**Business justification:** A managing partner hearing that certain groups seem less satisfied needs statistical evidence, not anecdote. Hypothesis testing converts observations into decision-quality statements.

### Hypothesis 1 - Do men and women report different job satisfaction?

**H0:** Mean satisfaction is equal between male and female respondents
**H1:** Mean satisfaction differs between male and female respondents
**Test:** Welch t-test | alpha = 0.05

```{r}
#| label: hypothesis1

h1_data <- df[df$gender %in% c("Male","Female"), ]
t_result <- t.test(satisfaction ~ gender, data = h1_data, var.equal = FALSE)
effect   <- cohens_d(satisfaction ~ gender, data = h1_data)

result_df <- data.frame(
  Statistic = c("Male Mean","Female Mean","t-statistic","p-value",
                "Cohen's d","Decision"),
  Value = c(
    round(mean(h1_data$satisfaction[h1_data$gender=="Male"]), 2),
    round(mean(h1_data$satisfaction[h1_data$gender=="Female"]), 2),
    round(t_result$statistic, 3),
    round(t_result$p.value, 4),
    round(abs(effect$Cohens_d), 3),
    ifelse(t_result$p.value < 0.05, "Reject H0", "Fail to reject H0")
  )
)
knitr::kable(result_df, caption = "Hypothesis 1 - Welch t-test Results")
```

```{r}
#| fig-height: 5

h1_data |>
  ggplot(aes(x = gender, y = satisfaction, fill = gender)) +
  geom_violin(alpha = 0.4, trim = FALSE) +
  geom_boxplot(width = 0.2, alpha = 0.9) +
  stat_summary(fun = mean, geom = "point", size = 4,
               colour = "#E8A020", shape = 18) +
  scale_fill_manual(values = c(Female = "#A23B72", Male = "#2E86AB"),
                    guide = "none") +
  scale_y_continuous(breaks = 1:10) +
  labs(title = "Hypothesis 1 - Job Satisfaction by Gender",
       subtitle = "Diamond = mean | Box = IQR | Violin = full distribution",
       x = NULL, y = "Job Satisfaction (1-10)")
```

### Hypothesis 2 - Does satisfaction differ across role types?

**H0:** Mean satisfaction is equal across all role categories
**H1:** At least one role category has a different mean satisfaction
**Test:** One-way ANOVA | alpha = 0.05

```{r}
#| label: hypothesis2

aov_result  <- aov(satisfaction ~ role, data = df)
aov_summary <- summary(aov_result)
eta         <- eta_squared(aov_result)

result2_df <- data.frame(
  Statistic = c("F-statistic","df between","df within","p-value",
                "Eta squared","Decision"),
  Value = c(
    round(aov_summary[[1]]$`F value`[1], 3),
    aov_summary[[1]]$Df[1],
    aov_summary[[1]]$Df[2],
    round(aov_summary[[1]]$`Pr(>F)`[1], 4),
    round(eta$Eta2[1], 3),
    ifelse(aov_summary[[1]]$`Pr(>F)`[1] < 0.05,
           "Reject H0 - satisfaction differs by role",
           "Fail to reject H0")
  )
)
knitr::kable(result2_df, caption = "Hypothesis 2 - One-Way ANOVA Results")
```

```{r}
#| fig-height: 6

df |> mutate(role = str_wrap(role, 20)) |>
  ggplot(aes(x = fct_reorder(role, satisfaction, median), y = satisfaction, fill = role)) +
  geom_boxplot(alpha = 0.8, show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "#E8A020") +
  scale_fill_manual(values = palette_cat) +
  coord_flip() + scale_y_continuous(breaks = 1:10) +
  labs(title = "Hypothesis 2 - Satisfaction by Role Type",
       subtitle = "Ordered by median | Diamond = mean",
       x = NULL, y = "Job Satisfaction (1-10)")
```

**Interpretation:** Satisfaction levels differ significantly across role types. The firm cannot apply a single HR policy uniformly - interventions must be role-specific, with administrative and IT staff receiving urgent attention.

---

## 8. Technique 4 - Correlation Analysis

**Business justification:** Before investing in HR initiatives, a managing partner needs to know which levers actually move satisfaction. Correlation identifies which variables move together.

```{r}
#| label: correlation
#| fig-height: 7
#| fig-width: 8

cor_data <- df[, c("satisfaction","work_life","performance","supervisor",
                   "compensation","hours_num","training_num","experience_num","tenure_num")]
cor_data <- cor_data[complete.cases(cor_data), ]

cor_labels <- c("Satisfaction","Work-Life","Performance","Supervisor",
                "Compensation","Hrs/Week","Training","Experience","Tenure")

cor_matrix <- cor(cor_data, method = "pearson")
colnames(cor_matrix) <- cor_labels
rownames(cor_matrix) <- cor_labels

corrplot(cor_matrix, method = "color", type = "upper", order = "hclust",
         col = colorRampPalette(c("#A23B72","white","#1B3A6B"))(200),
         addCoef.col = "black", number.cex = 0.8,
         tl.col = "#2C2C2C", tl.cex = 0.9, tl.srt = 45,
         title = "Pearson Correlation Matrix", mar = c(0,0,2,0))
```

**Top correlations and business implications:**

1. **Supervisor Support and Satisfaction (r = `r round(cor_matrix["Satisfaction","Supervisor"], 2)`)** - The strongest driver of satisfaction is feeling supported by managers. Investing in supervisor training has the highest return.

2. **Compensation and Satisfaction (r = `r round(cor_matrix["Satisfaction","Compensation"], 2)`)** - Compensation fairness is closely tied to satisfaction. Benchmarking salaries against market rates is a direct action.

3. **Work-Life Balance and Satisfaction (r = `r round(cor_matrix["Satisfaction","Work-Life"], 2)`)** - Long-hours cultures damage satisfaction. Flexible working policies could improve this.

---

## 9. Technique 5 - Logistic Regression

**Business justification:** Logistic regression predicts attrition risk, producing a model that can flag at-risk employees before they resign.

**Outcome:** leaving_binary (1 = any intention to leave, 0 = intends to stay)

```{r}
#| label: regression

log_model <- glm(
  leaving_binary ~ satisfaction + work_life + compensation +
                   supervisor + performance + hours_num +
                   training_num + tenure_num,
  data = df, family = binomial(link = "logit")
)

tidy_log <- tidy(log_model, conf.int = TRUE, exponentiate = TRUE)
tidy_log <- tidy_log[tidy_log$term != "(Intercept)", ]
tidy_log$term <- c("Job Satisfaction","Work-Life Balance","Compensation Fairness",
                   "Supervisor Support","Job Performance","Weekly Hours",
                   "Training Sessions","Tenure")
tidy_log$significant <- ifelse(tidy_log$p.value < 0.05, "Yes", "No")

knitr::kable(
  tidy_log[, c("term","estimate","conf.low","conf.high","p.value","significant")],
  col.names = c("Predictor","Odds Ratio","95% CI Low","95% CI High","p-value","Significant"),
  digits = 3,
  caption = "Logistic Regression - Predicting Attrition Intention"
)
```

```{r}
#| fig-height: 6

ggplot(tidy_log, aes(x = estimate, y = reorder(term, estimate),
                     colour = significant)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "#888888") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.25) +
  geom_point(size = 5) +
  scale_colour_manual(values = c(Yes = "#1B3A6B", No = "#AAAAAA"),
                      name = "Significant (p<0.05)") +
  labs(title = "Logistic Regression - Odds Ratios",
       subtitle = "Points right of dashed line increase attrition risk",
       x = "Odds Ratio", y = NULL)
```

```{r}
null_model   <- glm(leaving_binary ~ 1, data = df, family = binomial)
mcfadden_r2  <- round(as.numeric(1 - logLik(log_model)/logLik(null_model)), 3)
pred_class   <- ifelse(predict(log_model, type = "response") >= 0.5, 1L, 0L)
accuracy     <- round(mean(pred_class == df$leaving_binary, na.rm = TRUE), 3)

cat("McFadden R2:", mcfadden_r2, "| Accuracy:", accuracy, "| AIC:", round(AIC(log_model),1))
```

**Interpretation:** The model predicts attrition with `r round(accuracy*100,1)`% accuracy. Lower compensation fairness significantly increases attrition odds. Tenure is protective - longer-serving staff are less likely to leave. The firm should focus retention on staff in their first 1-3 years and conduct a compensation review immediately.

---

## 10. Integrated Findings

| Technique | Key Finding | Business Implication |
|-----------|-------------|---------------------|
| EDA | 73% express attrition intention | Immediate retention emergency |
| Visualisation | Admin and IT most at risk | Target these groups first |
| Hypothesis Testing | Satisfaction differs by role | Role-specific HR policies needed |
| Correlation | Supervisor support drives satisfaction | Invest in management training |
| Logistic Regression | Compensation and tenure predict attrition | Fix pay, focus on new staff |

**Single recommendation:** Conduct an immediate compensation benchmarking exercise, implement supervisor development programmes, and create formal retention plans for staff in their first three years.

---

## 11. Limitations & Further Work

1. **Convenience sampling** may over-represent certain demographics. A random probability sample would yield more generalisable findings.
2. **Self-report bias** may inflate performance and supervisor ratings.
3. **Cross-sectional design** captures one point in time. A longitudinal study tracking actual attrition would be more rigorous.
4. **Further work:** A machine learning classifier could generate individual-level attrition probability scores for proactive intervention.

---

## References

Adi, B. (2026). *AI-powered business analytics*. Lagos Business School. https://markanalytics.online

R Core Team. (2024). *R: A language and environment for statistical computing*. https://www.R-project.org/

Wickham, H. et al. (2019). Welcome to the tidyverse. *Journal of Open Source Software, 4*(43). https://doi.org/10.21105/joss.01686

---

## Appendix: AI Usage Statement

Claude (Anthropic) was used to assist with R code structure, visualisation syntax, and Quarto document formatting. All analytical decisions - technique selection, hypothesis formulation, variable choices, and business interpretations - were made independently by the author. The data collection instrument was designed and administered independently. All recommendations reflect the author's own analysis.
