---
title: "Workplace Experience & Retention Analytics in Nigerian Professional Services"
author: "Hannah Ishola-Oshodi"
date: today
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    toc-location: left
    code-fold: true
    code-summary: "Show Code"
    self-contained: true
    smooth-scroll: true
    highlight-style: github
    css: styles.css
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
  echo: true
---

```{r load-data}
library(tidyverse); library(ggplot2); library(scales); library(corrplot)
library(knitr); library(kableExtra); library(patchwork); library(broom)
library(car); library(effectsize); library(moments); library(GGally)
library(RColorBrewer); library(viridis); library(gridExtra); library(lubridate)
library(janitor); library(gt); library(gtExtras); library(glue)

lbs_colors <- c(primary="#1B3A6B", secondary="#E8A020", accent1="#2E86AB",
                accent2="#A23B72", accent3="#4CAF50", accent4="#FF6B35")
palette_cat <- c("#1B3A6B","#E8A020","#2E86AB","#A23B72","#4CAF50","#FF6B35")

theme_lbs <- function(base_size=12){
  theme_minimal(base_size=base_size) %+replace%
    theme(
      plot.title    = element_text(face="bold", size=base_size+3, colour="#1B3A6B", hjust=0, margin=margin(b=8)),
      plot.subtitle = element_text(size=base_size, colour="#555555", hjust=0, margin=margin(b=12)),
      plot.caption  = element_text(size=9, colour="#888888", hjust=1),
      axis.title    = element_text(face="bold", colour="#2C2C2C"),
      axis.text     = element_text(colour="#444444"),
      panel.grid.major = element_line(colour="#EBEBEB", linewidth=0.4),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill="white", colour=NA),
      plot.background  = element_rect(fill="white", colour=NA),
      legend.position  = "bottom",
      legend.title     = element_text(face="bold", size=10),
      strip.text       = element_text(face="bold", colour="#1B3A6B"),
      plot.margin      = margin(15,15,10,15)
    )
}
theme_set(theme_lbs())

raw <- read.csv("survey_data.csv", stringsAsFactors=FALSE, check.names=FALSE)

names(raw) <- c("timestamp","gender","education","experience","role",
                "firm_size","hours","training","satisfaction","work_life",
                "performance","supervisor","compensation","leaving",
                "leave_reason","recommend","tenure")

df <- raw
df[] <- lapply(df, function(x) if(is.character(x)) trimws(x) else x)
df[] <- lapply(df, function(x){ if(is.character(x)) gsub("\u2013", "-", x, fixed=FALSE) else x })

df$satisfaction  <- suppressWarnings(as.numeric(df$satisfaction))
df$work_life     <- suppressWarnings(as.numeric(df$work_life))
df$performance   <- suppressWarnings(as.numeric(df$performance))
df$supervisor    <- suppressWarnings(as.numeric(df$supervisor))
df$compensation  <- suppressWarnings(as.numeric(df$compensation))

df$performance[is.na(df$performance)] <- median(df$performance, na.rm=TRUE)
df$training[is.na(df$training) | df$training==""] <- "None"
df$experience[df$experience==""] <- NA
df <- df[!is.na(df$experience), ]

exp_levels  <- c("Less than 2 years","2-5 years","6-10 years","11-15 years","More than 15 years")
hr_levels   <- c("Less than 35","35-40","41-50","51-60","More than 60")
tr_levels   <- c("None","1-2","3-5","6-10","More than 10")
ten_levels  <- c("Less than 1 year","1-3 years","4-7 years","8-12 years","More than 12 years")
firm_levels <- c("1-20","21-50","51-100","101-250","More than 250")
leav_levels <- c("No, I intend to stay","Unsure","Yes, but not actively looking","Yes, actively looking")
rec_levels  <- c("Definitely not","Probably not","Not sure","Probably yes","Definitely yes")

df$experience <- factor(df$experience, levels=exp_levels, ordered=TRUE)
df$hours      <- factor(df$hours,      levels=hr_levels,  ordered=TRUE)
df$training   <- factor(df$training,   levels=tr_levels,  ordered=TRUE)
df$tenure     <- factor(df$tenure,     levels=ten_levels, ordered=TRUE)
df$firm_size  <- factor(df$firm_size,  levels=firm_levels,ordered=TRUE)
df$leaving    <- factor(df$leaving,    levels=leav_levels)
df$recommend  <- factor(df$recommend,  levels=rec_levels)
df$gender        <- factor(df$gender)
df$education     <- factor(df$education)
df$role          <- factor(df$role)
df$leave_reason  <- factor(df$leave_reason)

df$leaving_binary <- ifelse(df$leaving=="No, I intend to stay", 0L, 1L)
df$experience_num <- as.numeric(df$experience)
df$training_num   <- as.numeric(df$training)
df$hours_num      <- as.numeric(df$hours)
df$tenure_num     <- as.numeric(df$tenure)

n_obs <- nrow(df)
cat("Data loaded. N =", n_obs, "\n")
```

---

**Study Title:** Workplace Experience & Employee Retention Analytics in Nigerian Professional Services  
**Author:** Hannah Ishola-Oshodi | MBA Candidate, Lagos Business School  
**Date:** `r format(Sys.Date(), "%B %d, %Y")`  
**Observations:** `r n_obs` survey respondents | **Variables:** 16  
**Organisation:** Adeyemi-Akisanya Associates | TDSH - The Dispute Solutions Hub | Lagos, Nigeria

---

## 1. Executive Summary {#sec-executive-summary}

This study investigates workplace experience and employee retention intentions among professionals in Nigerian legal and professional services organisations. A structured online survey was administered to **`r n_obs` respondents** drawn from legal, finance, administrative, IT, and management roles across firms of varying sizes.

Five analytical techniques were applied: **Exploratory Data Analysis (EDA)** revealed that nearly 73% of respondents express some intention to leave their current organisation within 12 months. **Data Visualisation** exposed systematic patterns in satisfaction across role types and firm sizes. **Hypothesis Testing** confirmed statistically significant differences in job satisfaction across role categories. **Correlation Analysis** found that supervisor support and compensation fairness are the strongest predictors of overall satisfaction. **Logistic Regression** identified compensation fairness, work-life balance, and tenure as the most significant predictors of attrition intent.

**Key Recommendation:** Nigerian professional-services organisations should urgently review compensation structures and supervisory quality. A structured compensation benchmarking exercise and mandatory supervisor training programme, particularly targeting mid-sized firms (20-100 staff), would measurably reduce attrition risk.

---

## 2. Professional Disclosure {#sec-disclosure}

**Job Title:** Executive Assistant  
**Organisation:** Adeyemi-Akisanya Associates - Legal Practice & ADR Services, Lagos, Nigeria  
**Sector:** Legal Services, Workplace Mediation & Alternative Dispute Resolution

### Organisational Context

Adeyemi-Akisanya Associates is a Nigerian legal practice offering litigation, corporate advisory, and employment law services. Within the firm operates **TDSH - The Dispute Solutions Hub**, a specialist arm dedicated to workplace mediation and alternative dispute resolution (ADR) for organisations across Nigeria.

TDSH works with corporate clients to resolve workplace conflicts before they escalate into formal legal disputes. This requires a deep understanding of the human dynamics driving conflict - how employees feel about their compensation, how supported they feel by management, how fairly they perceive their working conditions, and whether they are considering leaving their organisation. These are early warning indicators of workplace dysfunction that, if left unaddressed, often result in grievances, industrial action, or litigation.

As Executive Assistant to the firm, I sit at the intersection of legal operations and people management. I coordinate internal administrative processes, support partner-level decision-making, and liaise with TDSH on client engagement logistics.

### Why This Research Matters to TDSH

The motivation for this study emerged directly from TDSH's operational needs. When organisations approach TDSH for mediation support, a recurring challenge is the absence of structured workforce data. This study demonstrates how a simple, anonymised workplace survey combined with rigorous data analytics can surface actionable intelligence for ADR practitioners.

**1 - Exploratory Data Analysis:** Before any mediation can begin, TDSH needs to understand the landscape - who are the employees, what do their satisfaction scores look like, where are the outliers. EDA provides this diagnostic overview, helping ADR practitioners identify which departments or role types require the most urgent attention.

**2 - Data Visualisation:** Legal professionals and senior managers are time-constrained. TDSH uses visual evidence to present findings to client leadership teams and to demonstrate the need for intervention.

**3 - Hypothesis Testing:** When a client claims that a problem is mainly in one department, TDSH needs to know whether these perceptions are statistically real or simply anecdotal. Hypothesis testing provides the rigour to answer these questions with confidence.

**4 - Correlation Analysis:** Understanding which workplace variables move together helps TDSH prioritise its recommendations. If supervisor support and job satisfaction are strongly correlated, then management training becomes the intervention of choice.

**5 - Logistic Regression:** A regression model that predicts which employees are at highest risk of leaving allows organisations to intervene before resignations occur and before disputes escalate. This predictive capability transforms TDSH from a reactive dispute resolver into a proactive workforce stability partner.

---

## 3. Data Collection & Sampling {#sec-data}

### 3.1 Source & Collection Method

Data were collected via a **structured online questionnaire** administered through Google Forms between **28 April 2026 and 1 May 2026**. The survey was designed by the researcher and disseminated through professional WhatsApp networks, direct email contact, and LinkedIn.

### 3.2 Sample Frame & Size

| Parameter | Detail |
|-----------|--------|
| Population | Professionals in Nigerian legal and professional-services firms |
| Sampling method | Purposive convenience sampling via professional networks |
| Sample size (collected) | 107 responses |
| Sample size (after cleaning) | `r n_obs` usable responses |
| Time period | 28 April - 1 May 2026 (4 days) |
| Platform | Google Forms (anonymous; no email addresses collected) |
| Distribution channel | WhatsApp, LinkedIn, direct email |

### 3.3 Ethical Statement

The survey was fully **anonymous** - no names, employee IDs, or email addresses were collected. A consent statement was included at the top of the Google Form making participation entirely voluntary.

### 3.4 Data Quality Notes

- **1 missing value** in `experience` - row excluded
- **1 missing value** in `performance` - replaced with column median  
- **19 missing values** in `training` - treated as "None"

---

## 4. Data Description {#sec-eda}

### 4.1 Dataset Overview

```{r data-overview}
summary_tbl <- df |>
  select(satisfaction, work_life, performance, supervisor, compensation) |>
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") |>
  group_by(Variable) |>
  summarise(
    N        = n(),
    Mean     = round(mean(Value, na.rm=TRUE), 2),
    Median   = median(Value, na.rm=TRUE),
    SD       = round(sd(Value, na.rm=TRUE), 2),
    Min      = min(Value, na.rm=TRUE),
    Max      = max(Value, na.rm=TRUE),
    Skewness = round(skewness(Value, na.rm=TRUE), 2)
  ) |>
  mutate(Variable = c("Compensation Fairness","Job Satisfaction","Job Performance","Supervisor Support","Work-Life Balance"))

summary_tbl |>
  gt() |>
  tab_header(
    title    = md("**Descriptive Statistics - Numeric Rating Variables (1-10 scale)**"),
    subtitle = md(glue::glue("Based on {n_obs} survey respondents, Nigerian Professional Services, April-May 2026"))
  ) |>
  fmt_number(columns = c(Mean, SD, Skewness), decimals = 2) |>
  cols_align(align = "center", columns = -Variable) |>
  tab_style(style = cell_fill(color = "#1B3A6B"), locations = cells_column_labels()) |>
  tab_style(style = cell_text(color = "white", weight = "bold"), locations = cells_column_labels()) |>
  tab_style(style = cell_fill(color = "#F4F6FB"), locations = cells_body(rows = seq(1, nrow(summary_tbl), 2))) |>
  opt_table_font(font = google_font("Inter")) |>
  tab_options(table.width = pct(100), table.font.size = 13)
```

### 4.2 Data Quality Issues Identified

**Issue 1 - Missing training responses (n = 19):** Nineteen respondents skipped the training question. These were imputed as "None" because the question asked for a count; skipping most plausibly indicates zero formal training.

**Issue 2 - Right-skewed satisfaction distribution:** Job satisfaction (skewness = `r round(skewness(df$satisfaction), 2)`) and performance scores are mildly negatively skewed, indicating ceiling effects. This is typical of self-report workplace surveys.

### 4.3 Variable Distributions

```{r distributions, fig.height=10, fig.width=10}
score_vars   <- c("satisfaction","work_life","performance","supervisor","compensation")
score_labels <- c("Job Satisfaction","Work-Life Balance","Job Performance","Supervisor Support","Compensation Fairness")

plots <- map2(score_vars, score_labels, function(var, label) {
  m <- mean(df[[var]], na.rm=TRUE)
  ggplot(df, aes(x = .data[[var]])) +
    geom_histogram(binwidth=1, fill=lbs_colors["primary"], colour="white", alpha=0.85) +
    geom_vline(xintercept=m, colour=lbs_colors["secondary"], linewidth=1.2, linetype="dashed") +
    annotate("text", x=m+0.4, y=Inf, vjust=2, label=paste0("Mean = ", round(m,1)),
             colour=lbs_colors["secondary"], fontface="bold", size=3.5) +
    scale_x_continuous(breaks=1:10) +
    labs(title=label, x="Score (1-10)", y="Count") +
    theme_lbs(base_size=11)
})

wrap_plots(plots, ncol=2) +
  plot_annotation(
    title    = "Distribution of Workplace Experience Scores",
    subtitle = "Dashed gold line = mean | Each plot = one of the five rating dimensions",
    theme    = theme_lbs()
  )
```

---

## 5. Technique 1 - Exploratory Data Analysis {#sec-eda2}

### 5.1 Business Justification

Before a law firm's managing partner can act on staff welfare data, they need to understand who is in the data, how scores are distributed, and whether any variables are suspect. EDA is the diagnostic that makes every subsequent analysis credible.

### 5.2 Demographic Profile

```{r demo-profile, fig.height=8, fig.width=10}
p1 <- df |> count(gender) |>
  ggplot(aes(x=fct_reorder(gender,n), y=n, fill=gender)) +
  geom_col(width=0.6, show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.2, fontface="bold", size=4) +
  scale_fill_manual(values=palette_cat) +
  coord_flip() + labs(title="Gender", x=NULL, y="Count") + ylim(0,55)

p2 <- df |> count(education) |>
  mutate(education=str_wrap(education,25)) |>
  ggplot(aes(x=fct_reorder(education,n), y=n, fill=education)) +
  geom_col(width=0.6, show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.2, fontface="bold", size=4) +
  scale_fill_manual(values=palette_cat) +
  coord_flip() + labs(title="Education Level", x=NULL, y="Count") + ylim(0,45)

p3 <- df |> count(role) |>
  mutate(role=str_wrap(role,22)) |>
  ggplot(aes(x=fct_reorder(role,n), y=n, fill=role)) +
  geom_col(width=0.6, show.legend=FALSE) +
  geom_text(aes(label=n), hjust=-0.2, fontface="bold", size=4) +
  scale_fill_manual(values=palette_cat) +
  coord_flip() + labs(title="Role Type", x=NULL, y="Count") + ylim(0,30)

p4 <- df |> count(leaving) |>
  ggplot(aes(x=fct_reorder(leaving,n), y=n, fill=leaving)) +
  geom_col(width=0.6, show.legend=FALSE) +
  geom_text(aes(label=paste0(n," (",round(n/n_obs*100),"%)")), hjust=-0.1, fontface="bold", size=3.8) +
  scale_fill_manual(values=c("#4CAF50","#E8A020","#FF6B35","#A23B72")) +
  coord_flip() + labs(title="Attrition Intention", x=NULL, y="Count") + ylim(0,40)

(p1+p2)/(p3+p4) +
  plot_annotation(
    title    = "Respondent Profile - 100 Professionals, Nigerian Legal & Professional Services",
    subtitle = "April-May 2026 Survey",
    theme    = theme_lbs()
  )
```

### 5.3 Outlier Detection

```{r outliers}
df |>
  select(satisfaction, work_life, performance, supervisor, compensation) |>
  pivot_longer(everything(), names_to="variable", values_to="score") |>
  mutate(variable = case_when(
    variable == "satisfaction" ~ "Job Satisfaction",
    variable == "work_life"    ~ "Work-Life Balance",
    variable == "performance"  ~ "Job Performance",
    variable == "supervisor"   ~ "Supervisor Support",
    variable == "compensation" ~ "Compensation Fairness",
    TRUE ~ variable)) |>
  ggplot(aes(x=variable, y=score, fill=variable)) +
  geom_boxplot(alpha=0.8, outlier.colour=lbs_colors["accent4"],
               outlier.size=3, outlier.shape=21, outlier.fill=lbs_colors["accent4"]) +
  scale_fill_manual(values=palette_cat, guide="none") +
  scale_y_continuous(breaks=1:10) +
  coord_flip() +
  labs(title="Boxplots of All Five Rating Dimensions",
       subtitle="Orange dots = statistical outliers | Scores on a 1-10 scale",
       x=NULL, y="Score")
```

### 5.4 Plain-Language Interpretation

The sample is evenly split between male and female professionals. **`r round(sum(df$leaving != "No, I intend to stay")/n_obs*100)`% express some level of intention to leave** - a significant talent-retention risk. Job satisfaction scores cluster around 5-8 out of 10, but compensation fairness sits noticeably lower, averaging `r round(mean(df$compensation), 1)` out of 10. Supervisor support scores are the most variable, indicating inconsistent management quality across the sample.

---

## 6. Technique 2 - Data Visualisation {#sec-viz}

### 6.1 Business Justification

A busy legal partner will not read a table of numbers. Visualisation translates statistical patterns into decisions. The five charts below tell a single, coherent story: who is most at risk of leaving, and why.

### 6.2 Visualisation Narrative - Five Charts, One Story

```{r viz1, fig.height=6}
df |>
  count(role, leaving) |>
  group_by(role) |> mutate(pct=n/sum(n)) |> ungroup() |>
  mutate(role=str_wrap(role,22)) |>
  ggplot(aes(x=role, y=pct, fill=leaving)) +
  geom_col(position="fill", width=0.65) +
  scale_fill_manual(values=c("#4CAF50","#A8C8E0","#E8A020","#A23B72"), name="Intention to Leave") +
  scale_y_continuous(labels=percent_format()) +
  labs(title="Chart 1 of 5 - Attrition Intention by Role Type",
       subtitle="Administrative and IT staff show the highest leaving intentions",
       x=NULL, y="Proportion of Respondents") +
  theme_lbs() + theme(axis.text.x=element_text(angle=30, hjust=1))
```

```{r viz2, fig.height=6}
df |>
  group_by(role) |>
  summarise(mean_sat=mean(satisfaction), se=sd(satisfaction)/sqrt(n()), .groups="drop") |>
  mutate(role=fct_reorder(str_wrap(role,25), mean_sat)) |>
  ggplot(aes(x=role, y=mean_sat)) +
  geom_segment(aes(xend=role, y=0, yend=mean_sat), colour=lbs_colors["accent1"], linewidth=1.2) +
  geom_point(size=6, colour=lbs_colors["primary"]) +
  geom_errorbar(aes(ymin=mean_sat-se, ymax=mean_sat+se), width=0.2, colour=lbs_colors["secondary"], linewidth=1) +
  geom_text(aes(label=round(mean_sat,1)), vjust=-1.2, fontface="bold", size=4, colour=lbs_colors["primary"]) +
  coord_flip() + scale_y_continuous(limits=c(0,10), breaks=0:10) +
  labs(title="Chart 2 of 5 - Mean Job Satisfaction by Role",
       subtitle="Error bars show +/-1 standard error | Legal/professional roles most satisfied",
       x=NULL, y="Mean Satisfaction Score (1-10)")
```

```{r viz3, fig.height=6}
df |>
  filter(leave_reason != "I am not considering leaving") |>
  count(leave_reason) |>
  mutate(pct=n/sum(n), leave_reason=fct_reorder(leave_reason,n)) |>
  ggplot(aes(x=leave_reason, y=pct, fill=leave_reason)) +
  geom_col(width=0.65, show.legend=FALSE) +
  geom_text(aes(label=paste0(n," (",percent(pct, accuracy=1),")")), hjust=-0.1, fontface="bold", size=4.2) +
  scale_fill_manual(values=palette_cat) +
  scale_y_continuous(labels=percent_format(), limits=c(0,0.55)) +
  coord_flip() +
  labs(title="Chart 3 of 5 - Why Are Professionals Considering Leaving?",
       subtitle="Among those considering departure | Better opportunities & compensation dominate",
       x=NULL, y="Share of Responses")
```

```{r viz4, fig.height=6}
ggplot(df, aes(x=compensation, y=satisfaction, colour=leaving)) +
  geom_jitter(size=3, alpha=0.7, width=0.15, height=0.15) +
  geom_smooth(aes(group=1), method="lm", se=TRUE, colour="#2C2C2C", fill="#DDDDDD", linewidth=1) +
  scale_colour_manual(values=c("#4CAF50","#E8A020","#FF6B35","#A23B72"), name="Intention to Leave") +
  scale_x_continuous(breaks=1:10) + scale_y_continuous(breaks=1:10) +
  labs(title="Chart 4 of 5 - Compensation Fairness vs Job Satisfaction",
       subtitle="Colour = attrition intention | Grey band = 95% confidence interval",
       x="Compensation Fairness Score (1-10)", y="Job Satisfaction Score (1-10)")
```

```{r viz5, fig.height=6}
df |>
  group_by(firm_size, role) |>
  summarise(mean_sat=mean(satisfaction), .groups="drop") |>
  mutate(role=str_wrap(role,18)) |>
  ggplot(aes(x=firm_size, y=role, fill=mean_sat)) +
  geom_tile(colour="white", linewidth=0.8) +
  geom_text(aes(label=round(mean_sat,1)), fontface="bold", colour="white", size=4.5) +
  scale_fill_gradientn(colours=c("#A23B72","#E8A020","#4CAF50"), limits=c(1,10), name="Mean\nSatisfaction") +
  labs(title="Chart 5 of 5 - Satisfaction Heatmap: Role x Firm Size",
       subtitle="Darker green = higher satisfaction | Darker red/purple = lower satisfaction",
       x="Firm Size (number of employees)", y=NULL) +
  theme_lbs() + theme(axis.text.x=element_text(angle=30, hjust=1))
```

**Integrated Story (Charts 1-5):** Administrative and IT staff show the highest attrition risk (Chart 1) and the lowest satisfaction scores (Chart 2). Their primary drivers are better opportunities and compensation (Chart 3). The strong positive relationship between compensation fairness and satisfaction (Chart 4) confirms compensation is a lever management can pull. Smaller firms (1-20 staff) consistently score lower on satisfaction across most roles (Chart 5).

---

## 7. Technique 3 - Hypothesis Testing {#sec-hypothesis}

### 7.1 Business Justification

A managing partner hearing that certain groups seem less satisfied needs statistical evidence, not anecdote. Hypothesis testing converts observations into decision-quality statements.

### 7.2 Hypothesis 1 - Do men and women report different job satisfaction?

**H0:** Mean job satisfaction is equal between male and female respondents  
**H1:** Mean job satisfaction differs between male and female respondents  
**Test:** Independent-samples Welch t-test | alpha = 0.05

```{r hypothesis1}
h1_data  <- df |> filter(gender %in% c("Male","Female"))
t_result <- t.test(satisfaction ~ gender, data=h1_data, var.equal=FALSE)
effect   <- cohens_d(satisfaction ~ gender, data=h1_data)

tibble(
  Statistic = c("Male Mean","Female Mean","t-statistic","Degrees of Freedom","p-value","Cohen's d","Interpretation"),
  Value = c(
    round(mean(h1_data$satisfaction[h1_data$gender=="Male"]),2),
    round(mean(h1_data$satisfaction[h1_data$gender=="Female"]),2),
    round(t_result$statistic,3),
    round(t_result$parameter,1),
    round(t_result$p.value,4),
    round(abs(effect$Cohens_d),3),
    ifelse(t_result$p.value < 0.05, "Reject H0 - significant difference", "Fail to reject H0 - no significant difference")
  )
) |>
  gt() |>
  tab_header(title=md("**Hypothesis 1 Results - Welch t-test**")) |>
  tab_style(style=cell_fill(color="#1B3A6B"), locations=cells_column_labels()) |>
  tab_style(style=cell_text(color="white", weight="bold"), locations=cells_column_labels()) |>
  tab_style(style=cell_fill(color=if(t_result$p.value < 0.05) "#E8F5E9" else "#FFF3E0"),
            locations=cells_body(rows=7)) |>
  opt_table_font(font=google_font("Inter")) |>
  tab_options(table.width=pct(80))
```

```{r h1-viz, fig.height=5}
h1_data |>
  ggplot(aes(x=gender, y=satisfaction, fill=gender)) +
  geom_violin(alpha=0.4, trim=FALSE) +
  geom_boxplot(width=0.2, alpha=0.9, outlier.shape=21, outlier.fill=lbs_colors["accent4"]) +
  stat_summary(fun=mean, geom="point", size=4, colour=lbs_colors["secondary"], shape=18) +
  scale_fill_manual(values=c(Female=lbs_colors["accent2"], Male=lbs_colors["accent1"]), guide="none") +
  scale_y_continuous(breaks=1:10) +
  labs(title="Hypothesis 1 - Job Satisfaction by Gender",
       subtitle="Diamond = mean | Box = IQR | Violin = full distribution",
       x=NULL, y="Job Satisfaction (1-10)")
```

### 7.3 Hypothesis 2 - Does job satisfaction differ across role types?

**H0:** Mean job satisfaction is equal across all role categories  
**H1:** At least one role category has a different mean satisfaction  
**Test:** One-way ANOVA | alpha = 0.05

```{r hypothesis2}
aov_result  <- aov(satisfaction ~ role, data=df)
aov_summary <- summary(aov_result)
eta         <- eta_squared(aov_result)

tibble(
  Statistic = c("F-statistic","df (between)","df (within)","p-value","Eta squared","Interpretation"),
  Value = c(
    round(aov_summary[[1]]$`F value`[1],3),
    aov_summary[[1]]$Df[1],
    aov_summary[[1]]$Df[2],
    round(aov_summary[[1]]$`Pr(>F)`[1],4),
    round(eta$Eta2[1],3),
    ifelse(aov_summary[[1]]$`Pr(>F)`[1] < 0.05, "Reject H0 - satisfaction differs by role", "Fail to reject H0")
  )
) |>
  gt() |>
  tab_header(title=md("**Hypothesis 2 Results - One-Way ANOVA**")) |>
  tab_style(style=cell_fill(color="#1B3A6B"), locations=cells_column_labels()) |>
  tab_style(style=cell_text(color="white", weight="bold"), locations=cells_column_labels()) |>
  opt_table_font(font=google_font("Inter")) |>
  tab_options(table.width=pct(80))
```

```{r h2-viz, fig.height=6}
df |>
  mutate(role=str_wrap(role,20)) |>
  ggplot(aes(x=fct_reorder(role, satisfaction, median), y=satisfaction, fill=role)) +
  geom_boxplot(alpha=0.8, show.legend=FALSE, outlier.colour=lbs_colors["accent4"]) +
  stat_summary(fun=mean, geom="point", shape=18, size=4, colour=lbs_colors["secondary"]) +
  scale_fill_manual(values=palette_cat) +
  coord_flip() + scale_y_continuous(breaks=1:10) +
  labs(title="Hypothesis 2 - Satisfaction by Role Type",
       subtitle="Ordered by median | Diamond = mean",
       x=NULL, y="Job Satisfaction (1-10)")
```

**Plain-language interpretation:** The statistical tests reveal `r ifelse(t_result$p.value < 0.05, "a significant difference in satisfaction between genders", "no significant gender difference in satisfaction")`. More importantly, satisfaction levels differ significantly across role types (ANOVA p = `r round(aov_summary[[1]]$"Pr(>F)"[1], 4)`). The firm cannot apply a single HR policy uniformly - interventions must be role-specific, with administrative and IT staff receiving the most urgent attention.

---

## 8. Technique 4 - Correlation Analysis {#sec-correlation}

### 8.1 Business Justification

Before investing in HR initiatives, a managing partner needs to know which levers actually move satisfaction and reduce attrition. Correlation analysis identifies which variables move together - guiding where limited budgets should be directed.

### 8.2 Correlation Matrix

```{r correlation, fig.height=7, fig.width=8}
cor_vars <- df |>
  select(satisfaction, work_life, performance, supervisor, compensation,
         hours_num, training_num, experience_num, tenure_num)

cor_labels <- c("Satisfaction","Work-Life","Performance","Supervisor",
                "Compensation","Hrs/Week","Training","Experience","Tenure")

cor_matrix <- cor(cor_vars, use="complete.obs", method="pearson")
colnames(cor_matrix) <- cor_labels
rownames(cor_matrix) <- cor_labels

corrplot(cor_matrix, method="color", type="upper", order="hclust",
         col=colorRampPalette(c("#A23B72","white","#1B3A6B"))(200),
         addCoef.col="black", number.cex=0.8,
         tl.col="#2C2C2C", tl.cex=0.9, tl.srt=45, cl.cex=0.8,
         title="Pearson Correlation Matrix - Workplace Experience Variables",
         mar=c(0,0,2,0))
```

### 8.3 Top 3 Correlations & Business Implications

```{r top-correlations}
cor_df <- as.data.frame(as.table(cor_matrix)) |>
  filter(as.integer(Var1) < as.integer(Var2)) |>
  rename(Variable1=Var1, Variable2=Var2, Correlation=Freq) |>
  arrange(desc(abs(Correlation))) |>
  head(5) |>
  mutate(Correlation=round(Correlation,3),
         Strength=case_when(abs(Correlation)>=0.7~"Strong", abs(Correlation)>=0.4~"Moderate", TRUE~"Weak"))

cor_df |>
  gt() |>
  tab_header(title=md("**Top 5 Correlations by Absolute Strength**")) |>
  fmt_number(columns=Correlation, decimals=3) |>
  data_color(columns=Correlation,
             fn=scales::col_numeric(palette=c("#A23B72","white","#1B3A6B"), domain=c(-1,1))) |>
  tab_style(style=cell_fill(color="#1B3A6B"), locations=cells_column_labels()) |>
  tab_style(style=cell_text(color="white", weight="bold"), locations=cells_column_labels()) |>
  opt_table_font(font=google_font("Inter")) |>
  tab_options(table.width=pct(90))
```

**Three strongest correlations and their business meaning:**

1. **Supervisor Support and Satisfaction (r = `r round(cor_matrix["Satisfaction","Supervisor"], 2)`):** The single strongest driver of satisfaction is how supported staff feel by their manager. Investing in supervisor development has the highest return.

2. **Compensation and Satisfaction (r = `r round(cor_matrix["Satisfaction","Compensation"], 2)`):** Compensation fairness is closely tied to satisfaction. Benchmarking salaries against market rates is a direct action.

3. **Work-Life Balance and Satisfaction (r = `r round(cor_matrix["Satisfaction","Work-Life"], 2)`):** Long-hours cultures in legal services directly damage satisfaction. Flexible working policies could shift this relationship positively.

---

## 9. Technique 5 - Logistic Regression {#sec-regression}

### 9.1 Business Justification

Logistic regression predicts attrition risk, producing a model that can flag at-risk individuals before they resign.

### 9.2 Model Specification

**Outcome variable:** leaving_binary (1 = any intention to leave, 0 = intends to stay)  
**Predictors:** satisfaction, work_life, compensation, supervisor, performance, hours_num, training_num, tenure_num

```{r logistic-regression}
log_model <- glm(
  leaving_binary ~ satisfaction + work_life + compensation +
                   supervisor + performance + hours_num + training_num + tenure_num,
  data=df, family=binomial(link="logit")
)

tidy_log <- tidy(log_model, conf.int=TRUE, exponentiate=TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(
    term = case_when(
      term == "satisfaction"  ~ "Job Satisfaction",
      term == "work_life"     ~ "Work-Life Balance",
      term == "compensation"  ~ "Compensation Fairness",
      term == "supervisor"    ~ "Supervisor Support",
      term == "performance"   ~ "Job Performance",
      term == "hours_num"     ~ "Weekly Hours Worked",
      term == "training_num"  ~ "Training Sessions",
      term == "tenure_num"    ~ "Tenure (years)",
      TRUE ~ term),
    significant = ifelse(p.value < 0.05, "Yes", "No"),
    p.value   = round(p.value, 4),
    estimate  = round(estimate, 3),
    conf.low  = round(conf.low, 3),
    conf.high = round(conf.high, 3)
  ) |>
  arrange(p.value)

tidy_log |>
  select(term, estimate, conf.low, conf.high, p.value, significant) |>
  rename(Predictor=term, `Odds Ratio`=estimate,
         `95% CI Lower`=conf.low, `95% CI Upper`=conf.high,
         `p-value`=p.value, Significance=significant) |>
  gt() |>
  tab_header(
    title    = md("**Logistic Regression - Predicting Attrition Intention**"),
    subtitle = md("Odds ratios > 1 increase attrition risk | Odds ratios < 1 are protective")
  ) |>
  fmt_number(columns=c(`Odds Ratio`,`95% CI Lower`,`95% CI Upper`), decimals=3) |>
  tab_style(style=cell_fill(color="#1B3A6B"), locations=cells_column_labels()) |>
  tab_style(style=cell_text(color="white", weight="bold"), locations=cells_column_labels()) |>
  opt_table_font(font=google_font("Inter")) |>
  tab_options(table.width=pct(100))
```

### 9.3 Coefficient Plot

```{r coef-plot, fig.height=6, fig.width=9, dev="png"}
ggplot(tidy_log, aes(x=estimate, y=reorder(term, estimate), colour=significant)) +
  geom_vline(xintercept=1, linetype="dashed", colour="#888888") +
  geom_errorbarh(aes(xmin=conf.low, xmax=conf.high), height=0.25, linewidth=1) +
  geom_point(size=5) +
  scale_colour_manual(values=c(Yes="#1B3A6B", No="#AAAAAA"), name="Significant (p<0.05)") +
  labs(title="Logistic Regression - Odds Ratios with 95% Confidence Intervals",
       subtitle="Points right of dashed line (OR > 1) increase attrition risk",
       x="Odds Ratio", y=NULL) +
  theme_lbs()
```

### 9.4 Model Diagnostics

```{r model-diagnostics}
null_model  <- glm(leaving_binary ~ 1, data=df, family=binomial)
mcfadden_r2 <- 1 - (logLik(log_model)/logLik(null_model))
pred_prob   <- predict(log_model, type="response")
pred_class  <- if_else(pred_prob >= 0.5, 1L, 0L)
accuracy    <- mean(pred_class == df$leaving_binary, na.rm=TRUE)

tibble(
  Metric = c("McFadden Pseudo R2","Model Accuracy","AIC","Observations"),
  Value  = c(round(as.numeric(mcfadden_r2),3), round(accuracy,3), round(AIC(log_model),1), n_obs)
) |>
  gt() |>
  tab_header(title=md("**Model Fit Statistics**")) |>
  tab_style(style=cell_fill(color="#1B3A6B"), locations=cells_column_labels()) |>
  tab_style(style=cell_text(color="white", weight="bold"), locations=cells_column_labels()) |>
  opt_table_font(font=google_font("Inter")) |>
  tab_options(table.width=pct(50))
```

**Plain-language interpretation:** The logistic regression model predicts attrition intention with approximately `r round(accuracy*100, 1)`% accuracy. Lower compensation fairness scores significantly increase the odds of attrition. Tenure is protective - longer-serving staff are less likely to leave. The firm should focus retention efforts on staff in their first 1-3 years and conduct a compensation review as the highest-priority intervention.

---

## 10. Integrated Findings {#sec-integrated}

The five analytical techniques applied in this study build progressively on one another to tell a single, coherent story about workplace experience in Nigerian professional services organisations.

| Technique | Key Finding | Business Implication for TDSH |
|-----------|-------------|-------------------------------|
| EDA | 73% express attrition intention; compensation rated lowest at mean 5.4/10 | Immediate retention emergency - scale of dissatisfaction justifies formal intervention |
| Visualisation | Administrative and IT staff show highest attrition risk; smaller firms score lowest | TDSH should target smaller legal firms as priority clients for proactive mediation |
| Hypothesis Testing | Satisfaction differs significantly across role types (ANOVA p < 0.05) | A single firm-wide HR policy is insufficient - role-specific interventions are needed |
| Correlation | Supervisor support and compensation fairness are strongest drivers of satisfaction | Management training and pay benchmarking are the two highest-return investments |
| Logistic Regression | Compensation fairness and tenure are strongest predictors of attrition (accuracy 77%) | Staff in first 1-3 years with low compensation scores are the highest-risk group |

### What This Means for TDSH

When TDSH engages with a client organisation experiencing workplace conflict, these findings provide a replicable analytical framework. A short anonymous survey can rapidly surface the key drivers of dissatisfaction in any organisation. The regression model demonstrates that attrition intent can be predicted with 77% accuracy from just eight observable variables.

**Single Overarching Recommendation:** Nigerian professional services organisations - particularly those with fewer than 50 staff - must urgently address compensation fairness and supervisory quality as the two primary levers of workforce stability. For TDSH, this translates into a clear consulting proposition: a structured workforce diagnostic, followed by targeted mediation and management development, can measurably reduce attrition risk and pre-empt workplace disputes before they escalate into formal legal proceedings.

---

## 11. Limitations & Further Work {#sec-limitations}

**1 - Convenience sampling bias:** Respondents were recruited through the researcher's professional WhatsApp networks and LinkedIn contacts. This approach may over-represent professionals in the researcher's immediate network, potentially limiting generalisability to the broader Nigerian legal and professional services sector.

**2 - Self-report bias:** All five rating variables are self-reported perceptions. Social desirability effects may inflate performance and supervisor ratings, while negativity bias may deflate compensation scores.

**3 - Cross-sectional design:** The survey captures a single snapshot in time. A longitudinal panel study tracking the same respondents over 12-18 months would allow true predictive validation of the regression model.

**4 - Missing training data:** Nineteen respondents did not answer the training question and were conservatively coded as having attended zero sessions.

**5 - Further work:** With a larger sample (n > 500), a machine learning classifier such as Random Forest or XGBoost could generate individual-level attrition probability scores. Combined with TDSH's mediation expertise, such a model could form the basis of a commercial workforce diagnostic product.

---

## GitHub Repository

[Click here to view GitHub Repository](https://github.com/hannahishola-hi/lbs-data-analytics){target="_blank"}

Source code and dataset available for reproducibility and transparency.

---

## References {#sec-references}

Adi, B. (2026). *AI-powered business analytics: A practical textbook for data-driven decision making - from data fundamentals to machine learning in Python and R*. Lagos Business School / markanalytics.online. https://markanalytics.online

R Core Team. (2024). *R: A language and environment for statistical computing* (Version 4.x). R Foundation for Statistical Computing. https://www.R-project.org/

Wickham, H. et al. (2019). Welcome to the tidyverse. *Journal of Open Source Software, 4*(43), 1686. https://doi.org/10.21105/joss.01686

Wickham, H. (2016). *ggplot2: Elegant graphics for data analysis*. Springer. https://doi.org/10.1007/978-3-319-24277-4

---

## Appendix: AI Usage Statement {#sec-ai}

This project was completed with assistance from Claude (Anthropic) as an AI coding assistant. Claude was used to help structure R code syntax, suggest appropriate visualisation approaches, and assist with YAML header configuration for the Quarto document. All analytical decisions - including the choice of techniques, the formulation of hypotheses, the selection of predictor variables for the regression model, and the business interpretation of all outputs - were made independently by the author based on course instruction and professional judgement. The data collection instrument was designed and administered independently by the author. All recommendations reflect the author's own analysis of findings.

---

*Data Analytics II - Case Study 1 | Lagos Business School | Prof Bongo Adi | April 2026*
